The paper on \Large \bf Blockchain Frontrunning

### 2018-09-13 frontrunning/Latex/sections/paper.tex
[frontrunning] notes from meeting sep 13, 2018

\subsection{What is Front-running?} %Definition of front-running
\label{sec:What is front-running?}

@@

%___________________________________________________________%
\section{Related Work}
As the traditional front-running was originated in trading financial instruments most of the literature are focused on financial aspects of the markets.  However, there are also a few policies regarding domain name registration front-running~\cite{sac022en33:online}

\subsection{Literature on Traditional Front-running}\label{traditionalFrontrunning}
As traditional front-running has been the subject of discussion for decades, there are many literatures on this subject. Here we focus on the historical context and how the regulations evolved to be what they are today.\\
Front-running was first appeared on the Chicago Board Options Exchange (\textit{CBoE})~\cite{markham1988front}, and was defined by \textit{Securities Exchange Commission} in 1977 as the following:
\begin{quote}
The practice of effecting an options transaction based upon non-public information regarding an impending block transaction\footnote{Block in the stock market is large number of shares, 10\,000 or more, to sell which will heavily changes the price} in the underlying stock, in order to obtain a profit when the options market adjusts to the price at which the block trades. ~\cite{sec1978optionsmarket}
\end{quote}

@@

@@

\section{Status ICO} 
As discussed in~\ref{sec:Front Running on the Blockchains}, capped ICOs are considerable applications to be investigated in terms of the front-running attacks, \textit{Status.im} ICO is such application. In June 2017, \textit{Status.im}~\cite{statuswhitepaper} started its ICO and reached the defined cap within 3 hours, resulting in close to 300,000 Ether of funds. In order to prevent few large investments to buy up all the tokens and limit the amount of Ether deposited in an investment, they used a \textit{fair} token distribution method named \textit{Dynamic Ceiling} as an attempt to increase the time window for smaller investors. They implemented multiple caps --ceiling-- in which, each had a maximum amount that could be deposited in. In this case, every deposit was checked by the smart contract and the exceeding amount was refunded to the sender while the accepted amount was sent to their multisignature wallet address ~\cite{statusicoanalysis}.

During the time the ICO was open for participation, there were reports of Ethereum network being unusuable and transactions were not confirming. Further study showed that some mining pools might have been manipulating the network for their own profit. In addition, there were many transactions sent with higher gas price to front-run other transactions, however these transactions were failing due to the restriction in the ICO smart contract to reject any transactions with higher than 50 \textit{GWei} gas price (as an attempt to prevent the known front-running behavior).

@@

As the result of this analysis, we discovered a wide range of blockchain front-running case studies that are yet to investigated. Therefore, in the next sections we dive deeper into blockchain specific front-running details and analyze top recent DApps for front-running attack vectors and possible solutions proposed to mitigate these issues.

@@

\subsection{Case Studies}  
As blockchain records are immutably recorded, there is enough historical data to analyze for possible front-running detection.There is no perfect criteria to list top DApps on Ethereum blockchain, for the purpose of this research we analyzed the top 25 DApps based on recent user activities\footnote{List of decentralized applications \url{https://dappradar.com/dapps}} in the first week of September 2018.

%ICO is only in use when the time window is happenig.  this is based on the last 24hours activities 
%justify why started with ICO and go to other apps

\subsubsection{Exchanges.} \label{sec:frontrunningExchanges}
Exchanges such as EtherDelta\footnote{also known as ForkDelta for the UI \url{https://forkdelta.app/}}, are closest form to decentral exchanges, however their orderbooks are stored on a central server they control and shown on their user interface (website). Other than the fact that they can re-order or censor the orders on their servers, there are other risks of front-running nature in these designs. Similarly 0xProtocol uses \textit{Relayers} which act as the orderbook holders and are not prune to cheating. One main issue with these exchanges is \textit{cancellation grief}. Although the orderbooks are not in the smart contract, in order to prevent denial of service attacks, user needs to send a transaction to cancel her order. In this case, when an adversarial actor sees the cancellation transaction, he sends a fill order transactions with higher gasPrice to get in front of the cancellation order and take the order before it is cancelled. As mentioned in ~\ref{minersfrontrunning}, miners are in a better control to perform such attack.

\subsubsection{Name Services.}
\noindent As mentioned in Section~\ref{sec:Front Running on the Blockchains}, although front-running attacks have been more showcased in the context of decentralized exchanges and trading systems, they are not yet limited to the financial markets. Front-running can also occur within other (non-financial) blockchain applications such as naming systems. Recently, many blockchain-based domain registrars have been designed and implemented to eliminate the role of central parties \ie domain name system (DNS) which itself introduces single point of failures in the entire web. Ethereum Name Service (ENS)\footnote{\url{https://ens.domains/}}, is the most popular decentral naming service that uses a sealed auction to facilitate the procedures of purchasing new \texttt{.eth} domain names. These domains which can be resolved using any Ethereum web3 compatible browser. Although a sealed auction (\textit{startAuctionsAndBid}) is used while implementing the ENS, the deposit should be equal or bigger than the bid, this information leakage could be used to front-run the transaction with the same bid amount. By doing so the front-runner invalidates the original bid and hence registers the domain name himself. Ghazal is another Ethereum-based naming system that allows users to register \texttt{.ghazal} domain names and bind certificates to those names~\cite{moosavighazal}.  In Ghazal model, a user would register domain name by executing the \textit{registerdomain} function from the Ghazal smart contract with the domain name in plain text as the function argument. As mentioned before, These transactions will sit in the mempool so that it would be mined by Ethereum miners and included in the block. During this period in which the transaction is not yet confirmed, front-running attack can occur by (i) dishonest miners and/or (ii) regular non-miner user. In the first case, a miner would intercept the \textit{registerdomain} transaction and register that name ahead of the user. A regular non-miner node in the Ethereum network can front-run other user's \textit{registerdomain} at a good profit by paying higher transaction fees. This practice allows the adversarial party  to register an unregistered domain immediately after someone searched for it, and sell the domain name to the users for a higher price. Note that, domain name front-running also occurs in the vanilla system. In this case, front-runners would monitor and identify domain names of interest, and sell it back to the other users with a higher price~\cite{sac022en33:online}. Front-running practice can happen in the traditional domain name system (DNS) as well. Domain name registrars can steal available domain names once seeing a query from users that looks up that domain's availability. In this case, the registrar can register the domain name ahead and sell it back to the user for a higher price.

%______________________________________________________________%
\subsubsection{Crypto-Collectibles Games.}

\subsubsection{Gambling.} 
Due to deterministic nature of blockchains, running gambling games based on randomness is close to impossible. There had been attempts to use block headers or such as the seed to randomness but almost all have failed as every data accessible to the smart contract at the time of execution, is also accessible to other smart contracts and they could predict the output on the run, thus prone to front-running and miner's manipulation. Many infamous gambling games are now running on Ethereum network and front-running them requires a level of sophistication and knowledge of the internal of the games and Ethereum.\\
One of the most popular smart contracts on Ethereum blockchain belongs to a controversial gambling game named \texttt{Fomo3d}, also known as \textit{Exit Scam}\footnote{\url{https://exitscam.me/play}}. The aim of this game is to be the last person purchasing the key when the time goes to zero in which every ticket purchase increases the timer by 30 seconds. It was believed that because of the nature of the game, it will never end, however on August 22, 2018 the first round of the game ended and the winner took 10,469 Ether\footnote{First winner of Fomo3D, won 10,469 Ether \url{ https://etherscan.io/tx/0xe08a519c03cb0aed0e04b33104112d65fa1d3a48cd3aeab65f047b2abce9d508}}, equivalent to 2.1 Million dollars on the time of writing. 
It is believed that the winner, used a sophisticated approach to finish last~\cite{fomo3dhacker}. The winner deployed a few high gas consumption smart contracts prior to the win, when the timer of the game reached about 3 minutes, the winner bought 1 ticket and sent multiple high gasPrice transactions to aforementioned smart contracts to lure the miners to mine these highly profitable transactions. Luckily enough the winner was able to congest the network in a way that no other ticket purchases were mined in the 3 minutes window and the miners were focused on the high gasPriced transactions. All the details mentioned here are drawn from the transaction history on the blockchain and unless otherwise proven seem to be the way the winner acted to win the pot. 
%probably needs more work to be good enough. more resources will be out in the coming days !

%______________________________________________________________%
\section{How to stop Front-running?} %We have have the applicability of the solutions inline.

>>>>

\subsection{Traditional Front-running} %Definition of front-running
\label{sec:What is front-running?}

@@

%\subsection{Literature on Traditional Front-running}\label{traditionalFrontrunning}
As traditional front-running has been the subject of discussion for decades, there are many literatures on this subject. Here we focus on the historical context and how the regulations evolved to be what they are today.\\
Front-running was first appeared on the Chicago Board Options Exchange (\textit{CBoE})~\cite{markham1988front}, and was defined by \textit{Securities Exchange Commission} in 1977 as the following:
\begin{quote}
The practice of effecting an options transaction based upon non-public information regarding an impending block transaction\footnote{Block in the stock market is large number of shares, 10\,000 or more, to sell which will heavily changes the price} in the underlying stock, in order to obtain a profit when the options market adjusts to the price at which the block trades. ~\cite{sec1978optionsmarket}
\end{quote}

@@

%
%___________________________________________________________%
%\section{Related Work}
%As the traditional front-running was originated in trading financial instruments most of the literature are focused on financial aspects of the markets.  However, there are also a few policies regarding domain name registration front-running~\cite{sac022en33:online}

%TODO: frontrunning categories: ICOs, non-ICOs  DApps. why we chose these, top recent activities.

@@

%TODO: but we missed ICOs becasue they are high intensity usage in short period of time and does not include in the list of the most used ones.
%

\section{Cases of Front-running in DApps}  
As blockchain records are immutably recorded, there is enough historical data to analyze for possible front-running detection.There is no perfect criteria to list top DApps on Ethereum blockchain, for the purpose of this research we analyzed the top 25 DApps based on recent user activities\footnote{List of decentralized applications \url{https://dappradar.com/dapps}} in the first week of September 2018.
%TODO:
%ICO is only in use when the time window is happenig.  this is based on the last 24hours activities 
%justify why started with ICO and go to other apps . refrence to ICO section for actual justification

% everything that has a line up , someone can jump in the line and front-run
% talk about velocity that seems to be not front-runnable but because of of the liquidity pool limitation, there can be an order that consumes the last chunk of liquidity "edge case" , to front run others and make their "option calls" fail

%Tragedy of the commons

\subsection{Markets and Exchanges.} \label{sec:frontrunningExchanges}
Exchanges such as EtherDelta\footnote{also known as ForkDelta for the UI \url{https://forkdelta.app/}}, are closest form to decentral exchanges, however their orderbooks are stored on a central server they control and shown on their user interface (website). Other than the fact that they can re-order or censor the orders on their servers, there are other risks of front-running nature in these designs. Similarly 0xProtocol uses \textit{Relayers} which act as the orderbook holders and are not prune to cheating. One main issue with these exchanges is \textit{cancellation grief}. Although the orderbooks are not in the smart contract, in order to prevent denial of service attacks, user needs to send a transaction to cancel her order. In this case, when an adversarial actor sees the cancellation transaction, he sends a fill order transactions with higher gasPrice to get in front of the cancellation order and take the order before it is cancelled. As mentioned in ~\ref{minersfrontrunning}, miners are in a better control to perform such attack.

\subsection{Name Services.}
\noindent As mentioned in Section~\ref{sec:Front Running on the Blockchains}, although front-running attacks have been more showcased in the context of decentralized exchanges and trading systems, they are not yet limited to the financial markets. Front-running can also occur within other (non-financial) blockchain applications such as naming systems. Recently, many blockchain-based domain registrars have been designed and implemented to eliminate the role of central parties \ie domain name system (DNS) which itself introduces single point of failures in the entire web. Ethereum Name Service (ENS)\footnote{\url{https://ens.domains/}}, is the most popular decentral naming service that uses a sealed auction to facilitate the procedures of purchasing new \texttt{.eth} domain names. These domains which can be resolved using any Ethereum web3 compatible browser. Although a sealed auction (\textit{startAuctionsAndBid}) is used while implementing the ENS, the deposit should be equal or bigger than the bid, this information leakage could be used to front-run the transaction with the same bid amount. By doing so the front-runner invalidates the original bid and hence registers the domain name himself. Ghazal is another Ethereum-based naming system that allows users to register \texttt{.ghazal} domain names and bind certificates to those names~\cite{moosavighazal}.  In Ghazal model, a user would register domain name by executing the \textit{registerdomain} function from the Ghazal smart contract with the domain name in plain text as the function argument. As mentioned before, These transactions will sit in the mempool so that it would be mined by Ethereum miners and included in the block. During this period in which the transaction is not yet confirmed, front-running attack can occur by (i) dishonest miners and/or (ii) regular non-miner user. In the first case, a miner would intercept the \textit{registerdomain} transaction and register that name ahead of the user. A regular non-miner node in the Ethereum network can front-run other user's \textit{registerdomain} at a good profit by paying higher transaction fees. This practice allows the adversarial party  to register an unregistered domain immediately after someone searched for it, and sell the domain name to the users for a higher price. Note that, domain name front-running also occurs in the vanilla system. In this case, front-runners would monitor and identify domain names of interest, and sell it back to the other users with a higher price~\cite{sac022en33:online}. Front-running practice can happen in the traditional domain name system (DNS) as well. Domain name registrars can steal available domain names once seeing a query from users that looks up that domain's availability. In this case, the registrar can register the domain name ahead and sell it back to the user for a higher price.

%______________________________________________________________%
\subsection{Crypto-Collectibles Games.}

\subsection{Gambling.} 
Due to deterministic nature of blockchains, running gambling games based on randomness is close to impossible. There had been attempts to use block headers or such as the seed to randomness but almost all have failed as every data accessible to the smart contract at the time of execution, is also accessible to other smart contracts and they could predict the output on the run, thus prone to front-running and miner's manipulation. Many infamous gambling games are now running on Ethereum network and front-running them requires a level of sophistication and knowledge of the internal of the games and Ethereum.\\
One of the most popular smart contracts on Ethereum blockchain belongs to a controversial gambling game named \texttt{Fomo3d}, also known as \textit{Exit Scam}\footnote{\url{https://exitscam.me/play}}. The aim of this game is to be the last person purchasing the key when the time goes to zero in which every ticket purchase increases the timer by 30 seconds. It was believed that because of the nature of the game, it will never end, however on August 22, 2018 the first round of the game ended and the winner took 10,469 Ether\footnote{First winner of Fomo3D, won 10,469 Ether \url{ https://etherscan.io/tx/0xe08a519c03cb0aed0e04b33104112d65fa1d3a48cd3aeab65f047b2abce9d508}}, equivalent to 2.1 Million dollars on the time of writing. 
It is believed that the winner, used a sophisticated approach to finish last~\cite{fomo3dhacker}. The winner deployed a few high gas consumption smart contracts prior to the win, when the timer of the game reached about 3 minutes, the winner bought 1 ticket and sent multiple high gasPrice transactions to aforementioned smart contracts to lure the miners to mine these highly profitable transactions. Luckily enough the winner was able to congest the network in a way that no other ticket purchases were mined in the 3 minutes window and the miners were focused on the high gasPriced transactions. All the details mentioned here are drawn from the transaction history on the blockchain and unless otherwise proven seem to be the way the winner acted to win the pot. 
%probably needs more work to be good enough. more resources will be out in the coming days !

\section{Cases of Front-running in ICOs} 
%icos are easy to be missed in the overall most used apps. 
%Make subsections here. one for ico's and why they are important and not in the most used daaps ones. one for  status ico, and different ones for each part
%justify why ICOs were missed in the previous section. argue why 100 other things like ICOs that were missing from the list . ICOs were missed becasue no ico was happening there, and also the website filteres through known applications, we could not think of any other scenarios that are similar to icos.

As discussed in~\ref{sec:Front Running on the Blockchains}, capped ICOs are considerable applications to be investigated in terms of the front-running attacks, \textit{Status.im} ICO is such application. In June 2017, \textit{Status.im}~\cite{statuswhitepaper} started its ICO and reached the defined cap within 3 hours, resulting in close to 300,000 Ether of funds. In order to prevent few large investments to buy up all the tokens and limit the amount of Ether deposited in an investment, they used a \textit{fair} token distribution method named \textit{Dynamic Ceiling} as an attempt to increase the time window for smaller investors. They implemented multiple caps --ceiling-- in which, each had a maximum amount that could be deposited in. In this case, every deposit was checked by the smart contract and the exceeding amount was refunded to the sender while the accepted amount was sent to their multisignature wallet address ~\cite{statusicoanalysis}.

During the time the ICO was open for participation, there were reports of Ethereum network being unusable and transactions were not confirming. Further study showed that some mining pools might have been manipulating the network for their own profit. In addition, there were many transactions sent with higher gas price to front-run other transactions, however these transactions were failing due to the restriction in the ICO smart contract to reject any transactions with higher than 50 \textit{GWei} gas price (as an attempt to prevent the known front-running behavior).

@@

%TODO:
%As the result of this analysis, we discovered a wide range of blockchain front-running case studies that are yet to investigated. Therefore, in the next sections we dive deeper into blockchain specific front-running details and analyze top recent DApps for front-running attack vectors and possible solutions proposed to mitigate these issues.

@@

\section{How to stop Front-running?} %We have have the applicability of the solutions inline.
